---
title: "Assignment 5"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Cleaning and Pre-Processing

## Loading Data and Basic Packages

### Load packages
First we need to load the basic packages for the manipulation of data. Other packages will be loaded later.
```{r packages}
library(tidyverse)
library(dplyr)
library(stringr)
library(arrow)
library(lubridate)
library(tsibble)
library(ggplot2)
```

### Load data
Now we load in the data. The app_data is the primary data we will use for now.
```{r loading}
# Personal level data from correction
person_level_data= read.csv('person_level_data.csv')
# Data of art unit changes
aus=read.csv("examiner_aus.csv")
# Get id crosswalk table
ids=read.csv("examiner_ids.csv")
# Get parquet data that details the transactionsfor each examiner, their gender and most likely ethnicity
App_data=read_parquet('apps_gender_rate.parquet')
```

## Clean Data
We have one Dataset that contains everything we might need. All the patent transactions, patent examiner ID, generd and ethnicity. We will need to remove the application status dates that are missing as we are relying on this field to predict the future.
```{r cleaning the data}
# Remove Nas
App_data <- App_data %>% 
  filter(!is.na(appl_status_date))

# Clean Date format
#get the date format cleaned
App_data$Date_time=as.POSIXct(App_data$appl_status_date, format="%d%b%Y")
#get the quarter number
App_data$Quarter_Year=as.character(yearquarter(App_data$Date_time))
#get the week number
App_data$Week_Year=as.character(yearweek(App_data$Date_time))

```

## Pre-process
Next we need to maks a few transformations to make sure that the data is in the format we need.First we need to filter the status update to a decison since we are trying to determine the production (number of application decisions by art unit in a week). Therefore we will remove all applications with the status "PEND". Since we know the only status tags are:
```{r Quick Check on disposal types}
unique(App_data$disposal_type)
```

```{r pre-process data 1}
#create new data fame with all manipulations. App_Data held as clean data
T_Data=App_data

#remove all values before 2018
T_Data <- T_Data %>% 
  filter(Date_time<= as.Date("2018-01-01"))



#Remove all the data we will not need based on application status
exclude_list=c("PEND")
T_Data <- T_Data %>%
  filter(!disposal_type %in% exclude_list)

#Remove all the date data that we will not need based on weeknum
exclude_list=c("PEND")
T_Data <- T_Data %>%
  filter(!disposal_type %in% exclude_list)

#Data Remain
nrow(T_Data)/nrow(App_data)*100

```
We therefore have 83.8% of our data remaining. This is an acceptable amout for out analysis.

Now we will need to group our data based on the processed applications. We will make a gender neutral and 2 gender inclusive(Male,Female) Data sets

```{r pre-process data 2}

#rename columns for prophet pacakge
#requires 'ds' and 'y'
#'ds' will be first day of week
#'y' will be the production value

GN_Data <- T_Data %>% 
  group_by(Week_Year) %>% 
  summarise(
    ds = min(Date_time),
    y=n(),
    )


G_Data <- T_Data %>% 
  group_by(Week_Year,gender) %>% 
  summarise(
    ds = min(Date_time), # I have repeatedly attempted to change this assumption but have been unable to
    y=n(),
    )


M_Data <- T_Data %>% 
  group_by(Week_Year,gender) %>% 
  summarise(
    ds = min(Date_time),
    y=n(),
    ) %>%
  filter(gender=="male")

F_Data <- T_Data %>% 
  group_by(Week_Year,gender) %>% 
  summarise(
    ds = min(Date_time),
    y=n(),
    ) %>%
  filter(gender=="female")


#remove(T_Data,App_data)

```
Now we need to separate the gender neutral dataframe to keep out the last year

```{r pre-process data 3}
fdate="2017-01-01"
GN_Data_Train <- GN_Data %>%
  filter(ds< as.Date(fdate))
GN_Data_Test <- GN_Data %>%
  filter(ds>= as.Date(fdate))

```


# Use 'Prophet to predict'

Now that we have our datasets we can use the prohet package to predict the rates and future state. Future is creating an empty dataframe for us for 365 days in the future.

```{r prediction 1}
#install.packages('prophet')
library(prophet)

m<-prophet(GN_Data_Train)
future <- make_future_dataframe(m, periods = 365)
tail(future)

```
Now we can make some predictions

```{r prediction 2}
forecast <- predict(m, future)
tail(forecast[c('ds', 'yhat', 'yhat_lower', 'yhat_upper')])

```

It looks like we are predicting between 3,600 and 4,300 applications per week in end of 2017. This is not accounting for hollidays, which we will need to fix.

```{r prediction 3}
plot(m, forecast)
```
We can look at the key feautres effecting the model
```{r prediction 4}
prophet_plot_components(m, forecast)

```
and then visualize the predictions!

```{r plot 2017 no holliday}
# Plot the prophet forecast with the test data points
ggplot(forecast, aes(ds, yhat)) +
  geom_ribbon(data = forecast,aes(ymin=yhat_lower,ymax=yhat_upper), alpha=0.1, color = "grey71")+
  geom_point(data = GN_Data_Train, aes(ds, y, color = "train data"), size = 2) +
  geom_point(data = GN_Data_Test, aes(ds, y, color = "test data"), size = 2) +
  geom_line(aes(color = "prophet forecast")) +
  scale_color_manual(values = c("blue", "red", "black"), labels = c("Forecast", "Test data", "Train data")) +
  xlab("Date") +
  ylab("y") +
  ggtitle("Forecast with Actual Data Points")
```

We can also take a look at the model evaluation

```{r prediction 2017 eval}
actuals <- GN_Data_Test$y
predictions <- tail(forecast$yhat,26)

MAE <- mean(abs(actuals - predictions))
MSE <- mean((actuals - predictions)^2)
RMSE <- sqrt(MSE)
MAPE <- mean(abs(actuals - predictions) / actuals)

print(MAE)
print(MSE)
print(RMSE)
print(MAPE)


```


## 2017 with Hollidays

The last prediction is not as clear as we could have hopped for. 

So let's account for US publich Hollidays based on: 
https://www.commerce.gov/hr/employees/leave/holidays
https://www.kaggle.com/datasets/donnetew/us-holiday-dates-2004-2021


```{r prediction 5}
Stat_Holliday <- data_frame(
  holiday = 'StatHolliday',
  ds = as.Date(c('2004-07-04','2005-07-04','2006-07-04','2007-07-04',
                 '2008-07-04','2009-07-04','2010-07-04','2011-07-04',
                 '2012-07-04','2013-07-04','2014-07-04','2015-07-04',
                 '2016-07-04','2017-07-04','2018-07-04','2019-07-04',
                 '2020-07-04','2021-07-04','2004-12-25','2005-12-25',
                 '2006-12-25','2007-12-25','2008-12-25','2009-12-25',
                 '2010-12-25','2011-12-25','2012-12-25','2013-12-25',
                 '2014-12-25','2015-12-25','2016-12-25','2017-12-25',
                 '2018-12-25','2019-12-25','2020-12-25','2021-12-25','2004-12-24','2005-12-24','2006-12-24','2007-12-24','2008-12-24','2009-12-24','2010-12-24','2011-12-24','2012-12-24','2013-12-24','2014-12-24','2015-12-24','2016-12-24','2017-12-24','2018-12-24','2019-12-24','2020-12-24','2021-12-24','2007-10-08','2012-10-08','2018-10-08','2006-10-09','2017-10-09','2005-10-10','2011-10-10','2016-10-10','2004-10-11','2010-10-11','2021-10-11','2009-10-12','2015-10-12','2020-10-12','2008-10-13','2014-10-13','2013-10-14','2019-10-14','2010-04-04','2007-04-08','2018-04-08','2004-04-11','2015-04-12','2012-04-15','2017-04-16','2020-04-19','2009-04-19','2014-04-20','2006-04-23','2011-04-24','2008-04-27','2019-04-28','2021-05-02','2005-05-01','2016-05-01','2013-05-05','2020-06-19','2021-06-19','2004-06-19','2005-06-19','2006-06-19','2007-06-19','2008-06-19','2009-06-19','2010-06-19','2011-06-19','2012-06-19','2013-06-19','2014-06-19','2015-06-19','2016-06-19','2017-06-19','2018-06-19','2019-06-19','2008-09-01','2014-09-01','2013-09-02','2019-09-02','2007-09-03','2012-09-03','2018-09-03','2006-09-04','2017-09-04','2005-09-05','2011-09-05','2016-09-05','2004-09-06','2010-09-06','2021-09-06','2009-09-07','2015-09-07','2020-09-07','2008-08-30','2014-08-30','2008-08-31','2014-08-31','2013-08-31','2019-08-31','2013-09-01','2019-09-01','2007-09-01','2012-09-01','2018-09-01','2007-09-02','2012-09-02','2018-09-02','2006-09-02','2017-09-02','2006-09-03','2017-09-03','2005-09-03','2011-09-03','2016-09-03','2005-09-04','2011-09-04','2016-09-04','2004-09-04','2010-09-04','2021-09-04','2004-09-05','2010-09-05','2021-09-05','2009-09-05','2015-09-05','2020-09-05','2009-09-06','2015-09-06','2020-09-06','2007-01-15','2018-01-15','2006-01-16','2012-01-16','2017-01-16','2005-01-17','2011-01-17','2010-01-18','2016-01-18','2021-01-18','2004-01-19','2009-01-19','2015-01-19','2014-01-20','2020-01-20','2008-01-21','2013-01-21','2019-01-21','2009-05-25','2015-05-25','2020-05-25','2008-05-26','2014-05-26','2013-05-27','2019-05-27','2007-05-28','2012-05-28','2018-05-28','2006-05-29','2017-05-29','2005-05-30','2011-05-30','2016-05-30','2004-05-31','2010-05-31','2021-05-31','2012-01-01','2013-01-01','2014-01-01','2015-01-01','2016-01-01','2017-01-01','2018-01-01','2019-01-01','2020-01-01','2011-01-01','2004-01-01','2005-01-01','2006-01-01','2007-01-01','2008-01-01','2009-01-01','2010-01-01','2021-01-01','2004-12-31','2005-12-31','2006-12-31','2007-12-31','2008-12-31','2009-12-31','2010-12-31','2011-12-31','2012-12-31','2013-12-31','2014-12-31','2015-12-31','2016-12-31','2017-12-31','2018-12-31','2019-12-31','2020-12-31','2021-12-31','2007-11-22','2012-11-22','2018-11-22','2006-11-23','2017-11-23','2005-11-24','2011-11-24','2016-11-24','2004-11-25','2010-11-25','2021-11-25','2009-11-26','2015-11-26','2020-11-26','2008-11-27','2014-11-27','2013-11-28','2019-11-28','2007-11-21','2012-11-21','2018-11-21','2006-11-22','2017-11-22','2005-11-23','2011-11-23','2016-11-23','2004-11-24','2010-11-24','2021-11-24','2009-11-25','2015-11-25','2020-11-25','2008-11-26','2014-11-26','2013-11-27','2019-11-27','2004-02-14','2005-02-14','2006-02-14','2007-02-14','2008-02-14','2009-02-14','2010-02-14','2011-02-14','2012-02-14','2013-02-14','2014-02-14','2015-02-14','2016-02-14','2017-02-14','2018-02-14','2019-02-14','2020-02-14','2021-02-14','2004-11-11','2005-11-11','2006-11-11','2007-11-11','2008-11-11','2009-11-11','2010-11-11','2011-11-11','2012-11-11','2013-11-11','2014-11-11','2015-11-11','2016-11-11','2017-11-11','2018-11-11','2019-11-11','2020-11-11','2021-11-11','2016-02-15','2015-02-16','2014-02-17','2020-02-17','2013-02-18','2019-02-18','2018-02-19','2012-02-20','2017-02-20','2011-02-21','2010-02-15','2021-02-15','2004-02-16','2009-02-16','2008-02-18','2007-02-19','2006-02-20','2005-02-21','2008-03-23','2005-03-27','2016-03-27','2013-03-31','2018-04-01','2010-04-04','2021-04-04','2015-04-05','2007-04-08','2012-04-08','2004-04-11','2009-04-12','2020-04-12','2006-04-16','2017-04-16','2014-04-20','2019-04-21','2011-04-24'))
,
  lower_window = 0,
  upper_window = 1
)

holidays <- bind_rows(Stat_Holliday)
```


```{r prediction 6}
m<-prophet(GN_Data_Train,holidays = holidays)

future <- make_future_dataframe(m, periods = 365)


forecast <- predict(m, future)
tail(future)

plot(m, forecast)
```

```{r plot the factors 2017 holliday}
prophet_plot_components(m, forecast)
```


```{r plot prediction 2017 holliday}
# Plot the prophet forecast with the test data points
ggplot(forecast, aes(ds, yhat)) +
  geom_ribbon(data = forecast,aes(ymin=yhat_lower,ymax=yhat_upper), alpha=0.1, color = "grey71")+
  geom_point(data = GN_Data_Train, aes(ds, y, color = "train data"), size = 2) +
  geom_point(data = GN_Data_Test, aes(ds, y, color = "test data"), size = 2) +
  geom_line(aes(color = "prophet forecast")) +
  scale_color_manual(values = c("blue", "red", "black"), labels = c("Forecast", "Test data", "Train data")) +
  xlab("Date") +
  ylab("y") +
  ggtitle("Forecast with Actual Data Points")

```


We can also take a look at the model evaluation

```{r prediction 2017 holliday eval}
actuals <- GN_Data_Test$y
predictions <- tail(forecast$yhat,26)

MAE <- mean(abs(actuals - predictions))
MSE <- mean((actuals - predictions)^2)
RMSE <- sqrt(MSE)
MAPE <- mean(abs(actuals - predictions) / actuals)

print(MAE)
print(MSE)
print(RMSE)
print(MAPE)


```
## 2016 data
Lets try removing 2017 data and predicting 2016 data instead

```{r prediction 2016}

fdate="2016-01-01"
edate="2017-01-01"
GN_Data_Train <- GN_Data %>%
  filter(ds< as.Date(fdate))

GN_Data_Test <- GN_Data %>%
  filter(ds>= as.Date(fdate))

GN_Data_Test <- GN_Data_Test %>%
  filter(ds<= as.Date(edate))

m<-prophet(GN_Data_Train)
future <- make_future_dataframe(m, periods = 365)
forecast <- predict(m, future)
plot(m,forecast)
```

```{r plot the factors 2016 holliday}
prophet_plot_components(m, forecast)
```



```{r plot prediction 2016 holliday}
ggplot(forecast, aes(ds, yhat)) +
  geom_ribbon(data = forecast,aes(ymin=yhat_lower,ymax=yhat_upper), alpha=0.1, color = "grey71")+
  geom_point(data = GN_Data_Train, aes(ds, y, color = "train data"), size = 2) +
  geom_point(data = GN_Data_Test, aes(ds, y, color = "test data"), size = 2) +
  geom_line(aes(color = "prophet forecast")) +
  scale_color_manual(values = c("blue", "red", "black"), labels = c("Forecast", "Test data", "Train data")) +
  xlab("Date") +
  ylab("y") +
  ggtitle("Forecast with Actual Data Points")


```

```{r model eval 2016}
actuals <- GN_Data_Test$y
predictions <- tail(forecast$yhat,52)

MAE <- mean(abs(actuals - predictions))
MSE <- mean((actuals - predictions)^2)
RMSE <- sqrt(MSE)
MAPE <- mean(abs(actuals - predictions) / actuals)


print(MAE)
print(MSE)
print(RMSE)
print(MAPE)

```




## 2016 data with hollidays
Lets try the same day with hollodays now


```{r prediction 2016 holliday}

m<-prophet(GN_Data_Train, holidays = holidays)
future <- make_future_dataframe(m, periods = 365)
forecast <- predict(m, future)

plot(m,forecast)

```


```{r plotting 2016 features}
prophet_plot_components(m, forecast)

```


```{r 2016 prediction plot}
ggplot(forecast, aes(ds, yhat)) +
  geom_ribbon(data = forecast,aes(ymin=yhat_lower,ymax=yhat_upper), alpha=0.1, color = "grey71")+
  geom_point(data = GN_Data_Train, aes(ds, y, color = "train data"), size = 2) +
  geom_point(data = GN_Data_Test, aes(ds, y, color = "test data"), size = 2) +
  geom_line(aes(color = "prophet forecast")) +
  scale_color_manual(values = c("blue", "red", "black"), labels = c("Forecast", "Test data", "Train data")) +
  xlab("Date") +
  ylab("y") +
  ggtitle("Forecast with Actual Data Points")

```
```{r model 2016 holliday stats}
actuals <- GN_Data_Test$y
predictions <- tail(forecast$yhat,52)

MAE <- mean(abs(actuals - predictions))
MSE <- mean((actuals - predictions)^2)
RMSE <- sqrt(MSE)
MAPE <- mean(abs(actuals - predictions) / actuals)

print(MAE)
print(MSE)
print(RMSE)
print(MAPE)

```

## Gender Data with 2016 Assumptions

Lets first clean up the data
```{r Gender }


fdate="2016-01-01"
edate="2017-01-01"
F_Data_Train <- F_Data %>%
  filter(ds< as.Date(fdate))

F_Data_Test <- F_Data %>%
  filter(ds>= as.Date(fdate))

F_Data_Test <- F_Data_Test %>%
  filter(ds<= as.Date(edate))


M_Data_Train <- M_Data %>%
  filter(ds< as.Date(fdate))

M_Data_Test <- M_Data %>%
  filter(ds>= as.Date(fdate))

M_Data_Test <- M_Data_Test %>%
  filter(ds<= as.Date(edate))




```


```{r Femal prediction}
m_F<-prophet(F_Data_Train,holidays=holidays)
future <- make_future_dataframe(m_F, periods = 365)
forecast_F <- predict(m_F, future)
plot(m_F,forecast_F)
```

```{r Male prediction}
m_M<-prophet(M_Data_Train ,holidays=holidays)
future <- make_future_dataframe(m_M, periods = 365)
forecast_M <- predict(m_M, future)
plot(m_M,forecast_M)
```

```{r prediction with gender}

ggplot(forecast, aes(ds, yhat)) +
  geom_ribbon(data = forecast_F,aes(ymin=yhat_lower,ymax=yhat_upper), alpha=0.4, color = "lightpink")+
  geom_ribbon(data = forecast_M,aes(ymin=yhat_lower,ymax=yhat_upper), alpha=0.4, color = "lightblue")+
  geom_point(data = F_Data_Train, aes(ds, y, color = "train data Femal"), size = 2, alpha=.2) +
  geom_point(data = F_Data_Test, aes(ds, y, color = "test data Female"), size = 2, alpha=.2) +
  geom_point(data = M_Data_Train, aes(ds, y, color = "train data Male"), size = 2,alpha=.2) +
  geom_point(data = M_Data_Test, aes(ds, y, color = "test data Male"), size = 2,alpha=.2) +
  scale_color_manual(values = c("magenta", "darkblue","pink2","lightblue3","black"), labels = c("Test Data Female", "Test data Male", "Training Data Female", "Test Data Female")) +
  xlab("Date") +
  ylab("y") +
  ggtitle("Forecast with Actual Data Points")

```

```{r model 2016 gender male satts}
actuals <- M_Data_Test$y
predictions <- tail(forecast_M$yhat,52)

MAE <- mean(abs(actuals - predictions))
MSE <- mean((actuals - predictions)^2)
RMSE <- sqrt(MSE)
MAPE <- mean(abs(actuals - predictions) / actuals)

print(MAE)
print(MSE)
print(RMSE)
print(MAPE)

```

```{r model 2016 gender female satts}
actuals <- F_Data_Test$y
predictions <- tail(forecast_F$yhat,52)

MAE <- mean(abs(actuals - predictions))
MSE <- mean((actuals - predictions)^2)
RMSE <- sqrt(MSE)
MAPE <- mean(abs(actuals - predictions) / actuals)

print(MAE)
print(MSE)
print(RMSE)
print(MAPE)

```
